Symmetric Peer-to-Peer Applications

### 2022-10-01 tpd-22.tex
- ate antes de 3.2

\IEEEtitleabstractindextext{%
\begin{abstract}
In multi-node collaborative applications, such as shared online documents,
users can interact remotely and yet share the exact same experience as if they
were in a single machine.
%
In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications} in which decentralized instances can broadcast asynchronous
events and yet conform to identical behavior.
%
Nodes are allowed to join and leave the network at any time.
Application developers must adhere to a restricted API, which is deterministic,
stateless, and only supports pre-allocated memory.
%
The middleware is responsible for synchronizing the events in a global shared
timeline across the network.
Based on memory snapshots and deterministic simulation, a ``time machine'' can
rollback conflicting peers to resynchronize their state.
%
TODO: application, results
\end{abstract}

@@

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Peers in the application form a dynamic graph network and communicate only
with direct neighbours, as a typical unstructured peer-to-peer network.
Events are flooded in the graph and are triggered locally with a small delay
to consider the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.
Our main contribution is the design of a middleware that runs on each peer to
ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

@@

Croquet%
\footnote{Croquet.io: https://croquet.io/}~\cite{croquet} guarantees
bit-identical real-time behavior for every user in a collaborative
distributed environment.
%
As detailed in Figure~\ref{fig.algos}, a centralized \emph{reflector} issues
periodic ticks, such that all nodes in the network advance together according
to a synchronized global shared clock.
If an user generates an event, it is delayed and sent back to the reflector,
which broadcasts the event in the next tick, which all nodes apply in sync
(including the originating node).
%
Croquet takes periodic snapshots of the whole application state in order to
support late joins to a running session.
The new node just needs to restore the latest snapshot and simulate the
remaining events to reach the current running state.
%
As indicated in Figure~\ref{fig.table}, Croquet relies on a centralized
network, in which nodes advance in sync, and in which event outcomes depend
on the central server to be online.

GALS~\cite{gals} shares the same goals with Croquet, but with some tradeoffs,
mostly favoring clients with slow connections.
As detailed in Figure~\ref{fig.algos}, instead of advancing ticks in sync with
the server, clients have their own local clocks.
Also, event generation is delayed dynamically according to the slowest client.
%
On the one hand, ticks do not generate any traffic and clients have smooth
frame transitions, even those with poor connections.
On the other hand, event generation requires two round trips to the server and
clients may experience occasional freezes.
The syncing protocol also requires bookkeeping to deal with clock drifts and
client disconnections~\cite{gals}.
%
As indicated in Figure~\ref{fig.table}, GALS relies on a centralized
network, in which nodes advance independently, and in which event outcomes
depend on the central server to be online.

In both solutions, the network advances together as a whole in real time, with
a total order among events, which is determined by a central server that must
be permanently online.
All clients compute the events in sequence, respecting their timestamps, and
using only deterministic and stateless calls.
This way, it is guaranteed that all clients reach the same state, also going
through bit-identical steps.

An antagonistic approach to deterministic reactions to events is to model the
application with conflict-free replicated data types (CRDTs)~\cite{crdts}.
CRDTs are designed in such a way that all operations are commutative, so that
the order in which they are applied does not affect the final outcome.
%
As detailed in Figure~\ref{fig.algos}, peers can communicate operations
directly to each other with no central authority.
Also, since operations need not to be ordered, they can be applied at the
very first moment they are generated or received, even if the peers are
offline.
%
Since operations must be commutative, it is not possible to timestamp them
in a global shared clock.
Hence, there is no notion of a timeline in which peers go through
bit-identical steps, but they do eventually reach the same state.
%
On the one hand, the main advantage of CRDTs is the support for local-first
software~\cite{local}, since they can work offline in the same way as online.
On the other hand, they provide very restrictive APIs (the CRDTs themselves),
and do not support real-time consensus among peers.
%
As indicated in Figure~\ref{fig.table}, CRDTs work on decentralized networks,
in which peers advance independently, and in which operations can be applied
immediately, even while offline.
%
Automerge along with its accompanying Hypermerge peer-to-peer infrastructure
is an example of an implementation of decentralized
CRDTs~\cite{p2p.automerge,p2p.pushpin}.

@@

In this section, we describe the architecture of our middleware and also the
API in which developers have to conform with.
%
The middleware is written in C and relies on the library \emph{SDL}%
\footnote{SDL: \url{www.libsdl.org}}
to deal with the network and input events, such as timers, mouse clicks and
key presses.
The full source code%
\footnote{TODO}
is under 500 lines of code.

@@

We guide our discussion through the didactic example of Figure~\ref{fig.move},
in which multiple users use the arrow keys to control the direction of a
moving rectangle.
The six instances are connected in a peer-to-peer network, such that some
peers can only communicate indirectly.
%
The application is symmetric in the sense that if any peer presses a key, a
corresponding event is broadcast and applied in all peers simultaneously, as
if they were a single mirrored instance.
At any moment, if a peer pauses the application, all peers are guaranteed to
pause exactly at the same frame with the rectangle being ``bit-identically'' at
the same position.

@@

In Figure~\ref{fig.middleware}, a peer is depicted as an application that uses
an API to communicate with the middleware.
The application itself is not aware of the network and has no direct access to
the other peers, which communicate transparently through the middleware.
The application generates asynchronous events that go through the middleware,
which synchronizes them with a timestamp in the future, such that all peers can
agree.
The middleware controls the execution of the application by issuing ticks and
synchronized events.

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
As described in the Introduction, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate delay so that they reach other peers in time to be applied in sync
in a common deadline.
%We first assume a predetermined delay for the sake of simplicity.
However, a peer that is locally ahead of time may receive an event that should
have been applied in the past, event considering the delay.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
As we detail further, rolling back a peer requires to simulate its whole
execution from the beginning (or from a snapshot in between).
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because:
    (a) applications are launched locally at different times, and
    (b) internal clocks may diverge over time (e.g., from rollbacks or timer
        polling inaccuracies).
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
\item \textbf{Peer Churn:}
Considering that we target peer-to-peer networks, it is important that the
middleware can handle arrival and departure of peers seamlessly.
Note that high churn may lead to intermittent network partitions.
Given our unstructured approach, nothing needs to be done when a peer leaves
the network.
However, when a peer joins the network, the middleware just needs to ensure
that it receives all events in order, which is trivial since they all hold an
absolute timestamp.
\end{enumerate}

@@

The middleware expects to take full control of the application execution in
order to control its timeline and disseminate events in the network.
For this reason, the basic programming API is to call a \code{loop} function
passing a set of callbacks, which the middleware uses to communicate with the
application:

@@

The middleware assumes that each peer assigns to itself a unique identifier in
a contiguous range.
The FPS must be the same in all peers to ensure that the simulation executes
exactly the same.

@@

The global variable \code{G} must hold the full application state.
This allows the middleware to take memory snapshots and rollback to a previous
state.
If dynamic allocation is required, it is necessary to hold a finite memory
pool in \code{G} with a custom allocator.
For instance, the moving rectangle application only requires to hold the
current position and direction:

@@

The application generates events that need to be broadcast to the network so
that all peers behave the same.
In this sense, application events may differ from low-level local system
events, which may or may not be mapped 1-to-1 in the callback \code{cb\_evt}.

There middleware predefines two events for all applications:
    \code{P2P\_EVT\_START} represents the beginning of the simulation, and
    \code{P2P\_EVT\_TICK} called at every frame.
The application-specific events must be declared by the programmer in an
enumeration.
In our application, we define a new event \code{P2P\_EVT\_KEY} to represent key
presses:

@@

The middleware calls \code{cb\_ini} twice: at the beginning and at the end of
the loop.
It should be used to initialize (and finalize) the network topology, as well
as static globals that can live outside the simulation memory, such as the
window and image textures:

@@

The middleware calls \code{cb\_sim} on every frame, passing an application
event, to affect the simulation state \code{G} deterministically.
The callback must never perform any side effects, such as stateful calls or
rendering frames.
In our example, we modify the state of the rectangle as follows:
    (a) reset its position and speed on \code{START},
    (b) increment its position based on the current speed on \code{TICK}, and
    (c) modify its speed based on \code{KEY}:

@@

The calls to \code{cb\_sim} normally happen during real-time simulation, i.e.,
while the user is interacting with the application.
However, if the peer is out of sync, the middleware may ``time travel'' and
call the simulation many times to go back and forth and catch up with real
time.
This is the reason why we define \code{START} as an event like any other, since
travelling requires to simulate the execution from the beginning, event by
event.

@@

Time travelling is also the reason why \code{cb\_sim} must be split from the
rendering callback \code{cb\_ren}~\cite{tml.js}, which would otherwise render
the screen multiple times while travelling.
%, which the middleware calls only on real-time frames.
In our example, \code{cb\_ren} just needs to redraw the rectangle at the
current position:

{\footnotesize
\begin{verbatim}
void cb_eff (void) {
    <...> // clears the SDL window
    // redraws the rectangle at the current position
    SDL_Rect r = { G.x, G.y, 20, 20 };
    SDL_DrawRect(&r);
}
\end{verbatim}
}

@@

The \code{cb\_evt} callback is called by the middleware in real time, whenever
a local SDL event occurs.
This callback has two goals:
    (a) map and broadcast application events, and
    (b) provide instant feedback to the user.
As mentioned previously, not all local low-level events may generate
application events to be broadcast in the network.
The callback is allowed to perform side effects, such as modifying the network
topology, or exhibiting some feedback on the screen.
In our example, we only generate application events for the 5 key events, and
we also exhibit the pressed key in real time:

@@

Regarding (a), if a peer experiences a time travel, it means that it received
an event after its deadline.
We count all time travel occurrences during the simulation.
%
Regarding (b), we use an absolute simulation clock to track, in real time, the
execution pace of each peer in comparison to a reference peer most advanced in
time.
%
Regarding (c), we designate some peers to leave and join the network
periodically, possibly creating partitions, to assert that they all catch up in
real time.

@@

Figure~\ref{TODO} results...
- colors indicate feasibility
%
We now detail how we instrument the middleware implementation to measure the
properties above:
%
\begin{enumerate}
%
\item \textbf{Recurrence of time travels:}
Every time a peer is ahead of time and needs to travel back, we output its id
and how much ticks it needs to travel back.
At the end, we measure the percentage of the sum of all time travel ticks over
the total simulation ticks (\code{TODO} in Figure~\ref{TODO}).
As an example, if all peers sum $1000$ time travel ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
\item \textbf{Real-time pace of peers:}
Every time a peer receives \code{SYNC} packets and is behind, we output its id
and how much ticks it needs to fast forward.
At the end, we measure the percentage of the sum of all forward ticks over the
total simulation ticks (\code{TODO} in Figure~\ref{TODO}).
As an example, if all peers sum $1000$ forward ticks over $100k$ accumulated
simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a time
mismatch.
%
Each peer outputs its current tick continuously, allowing us to compare them
in the shared timeline.
Each time we see a greater tick never seen in the timeline before, we count
smaller ticks from other peers (considering latency), which would constitute
time violations (\code{TODO} in Figure~\ref{TODO}).
We expect this counter to be close to zero even considering the whole
simulation.
%
\item \textbf{Correctness of unstable peers:}
- state (even for stable peers)
- time
- remove 13 (nao partition) e 10 (partition)
- qto tempo leva para sync
%
\end{enumerate}

>>>>

\IEEEtitleabstractindextext{%
\begin{abstract}
In collaborative networked applications, such as shared online documents,
users can interact remotely and yet share the exact same experience as if they
were in a single machine.
%
In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications} in which decentralized instances can broadcast asynchronous
events and yet conform to identical behavior.
%
Peers are allowed to join and leave the network at any time.
Application developers must adhere to a restricted API, which is deterministic,
stateless, and only supports pre-allocated memory.
%
The middleware is responsible for synchronizing the events in a global shared
timeline across the network.
Also, ased on memory snapshots and deterministic simulation, a ``time machine''
can rollback conflicting peers to resynchronize their state.
%
TODO: application, results
\end{abstract}

@@

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Peers in the application form a dynamic network graph and communicate only
with direct neighbours, as in typical unstructured peer-to-peer networks.
Events are flooded in the graph and are triggered locally with a small delay
to compensate the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.
Our main contribution is the design of a middleware that runs on each peer to
ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

@@

Croquet%
\footnote{Croquet.io: https://croquet.io/}~\cite{croquet} guarantees
bit-identical real-time behavior for every user in a collaborative
distributed environment.
%
As detailed in Figure~\ref{fig.algos}, a centralized \emph{reflector} issues
periodic ticks, such that all nodes in the network advance together according
to a synchronized global shared clock.
If a user generates an event, it is delayed and sent back to the reflector,
which broadcasts the event in the next tick, which all nodes apply in sync.
%
Croquet takes periodic snapshots of the whole application state in order to
support late joins to a running session.
The new node just needs to restore the latest snapshot and simulate the
remaining events to reach the current running state.
%
As indicated in Figure~\ref{fig.table}, Croquet relies on a centralized
network, in which nodes advance in sync, and in which event outcomes depend
on the central server to be online.

GALS~\cite{gals} shares the same goals with Croquet, but with some tradeoffs,
mostly favoring clients with slow connections.
As detailed in Figure~\ref{fig.algos}, instead of advancing ticks in sync with
the server, clients have their own independent local clocks.
Event generation requires two round trips (\emph{when→ask→ack→tick}) and is
delayed dynamically according to the slowest client.
%
On the one hand, ticks do not generate any traffic and clients have smooth
frame transitions, even those with poor connections.
On the other hand, events are slower and clients may experience occasional
freezes.
The syncing protocol also requires bookkeeping to deal with clock drifts and
client disconnections~\cite{gals}.
%
As indicated in Figure~\ref{fig.table}, GALS also relies on a centralized
network, but in which nodes advance time independently, and in which event
outcomes still depend on the central server to be online.

In both solutions, the network advances together as a whole in real time, with
a total order among events, which is determined by a central server that must
be permanently online.
All clients compute the events in sequence, respecting their timestamps, and
using only deterministic and stateless calls.
This way, it is guaranteed that all clients reach the same state, and
observe bit-identical steps.

An antagonistic approach to deterministic reactions to events is to model the
application with conflict-free replicated data types (CRDTs)~\cite{crdts}.
CRDTs are designed in such a way that all operations are commutative, so that
the order in which they are applied does not affect the final outcome.
%
As detailed in Figure~\ref{fig.algos}, peers can communicate operations
directly to each other with no central authority.
Also, since operations need not to be ordered, they can be applied at the
very first moment they are generated or received, even if the peers are
offline.
%
Since operations must be commutative, it is not possible to timestamp them
in a unique global shared clock.
Hence, there is no notion of a timeline in which peers go through
bit-identical steps, but they do eventually reach the same state.
%
On the one hand, the main advantage of CRDTs is to support local-first
software~\cite{local}, since they can work offline in the same way as online.
On the other hand, they provide very restrictive APIs (the CRDTs themselves),
and do not support real-time consensus among peers.
%
As indicated in Figure~\ref{fig.table}, CRDTs work on decentralized networks,
in which peers advance independently, and in which operations can be applied
immediately, even while offline.
%
Automerge along with its accompanying Hypermerge peer-to-peer infrastructure
is an example of an implementation of decentralized
CRDTs~\cite{p2p.automerge,p2p.pushpin}.

@@

In this section, we describe the architecture of our middleware and also the
API in which developers have to conform with.
%
The middleware is written in C and relies on the library \emph{SDL}%
\footnote{SDL: \url{www.libsdl.org}}
to deal with the network, timers, and input events, such as mouse clicks and
key presses.
The full source code%
\footnote{TODO}
is under 500 lines of code.

@@

We guide our discussion through the didactic example of Figure~\ref{fig.move},
in which multiple users use the arrow keys to control the direction of a
moving rectangle.
The six instances are connected in a peer-to-peer network, such that some
peers can only communicate indirectly.
%
The application is symmetric in the sense that if any user presses a key, a
corresponding event is broadcast and applied in all peers simultaneously, as
if they were a single mirrored machine.
At any moment, if a user pauses the application, all peers are guaranteed to
pause exactly at the same frame with the rectangle being ``bit-identically'' at
the same position.

@@

In Figure~\ref{fig.middleware}, a peer is depicted as an application that uses
an API to communicate with the middleware.
The application itself is not aware of the network and has no direct access to
the other peers, which communicate transparently through the middleware.
The application generates asynchronous events that go through the middleware,
which synchronizes them with a timestamp in the future, such that all peers can
satisfy.
The middleware controls the execution of the application by issuing ticks and
synchronized events.

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
As described in the Introduction, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate delay so that they can reach other peers in time to be applied in
sync.
%We first assume a predetermined delay for the sake of simplicity.
Nevertheless, a peer that is locally ahead of time may receive an event that
should have been applied in the past, even considering the delay.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
As we detail further, rolling back a peer requires to simulate the whole
application execution from the beginning.
As an optimization, the middleware takes periodic snapshots locally to avoid
full simulation.
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because applications are launched locally at different times, and
because internal clocks may diverge over time (e.g., from rollbacks or timer
polling inaccuracies).
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
\item \textbf{Peer Churn:}
Considering that we target peer-to-peer networks, it is important that the
middleware can handle arrival and departure of peers seamlessly.
Note that high churn may lead to intermittent network partitions.
Given our unstructured approach, nothing needs to be done when a peer leaves
the network.
However, when a peer joins the network, the middleware needs to ensure that the
peer receives all events that ever happened in order, which is trivial since
they all hold an absolute timestamp in the shared timeline.
\end{enumerate}

@@

The middleware expects to take full control of the application execution in
order to manipulate its timeline and disseminate events in the network.
For this reason, the basic programming API is to call a \code{loop} function
passing a set of callbacks, which the middleware uses to communicate with the
application:

@@

The middleware assumes that each peer assigns itself a unique identifier in a
contiguous range.
The FPS must be the same in all peers to ensure bit-identical simulation.

@@

The variable \code{G} passed to the middleware holds the full application
state.
This allows the middleware to take memory snapshots and rollback to previous
states.
If dynamic allocation is required, it is necessary to hold a finite memory
pool in \code{G} with a custom allocator.
In our guiding example of Figure~\ref{fig.move}, the moving rectangle
application only requires to hold the current position and direction:

@@

The application generates events that need to be broadcast to the network so
that all peers behave the same.
Note that the application events may (or may not) differ from low-level local
system events.
%which may or may not be mapped 1-to-1 in the callback \code{cb\_evt}.
%
The middleware predefines two events for all applications:
    \code{P2P\_EVT\_START} represents the beginning of the simulation, and
    \code{P2P\_EVT\_TICK} is generated every frame.
The application-specific events must be declared by the programmer in an
enumeration.
In our application, we define a new event \code{P2P\_EVT\_KEY} to represent key
presses:

@@

The middleware calls \code{cb\_ini} twice: at the beginning and at the end of
the loop.
The callback initializes (and finalizes) the network topology, as well as
static immutable globals that can live outside the simulation memory, such as
the window and image textures in our example:

@@

The middleware calls \code{cb\_sim} every frame, which affects the simulation
state \code{G} deterministically.
The occurring event is passed as an argument to the callback, which must never
perform any side effects, such as stateful calls or rendering frames.
In our example, we modify the state of the rectangle as follows:
    (a) reset its position and speed on \code{START},
    (b) increment its position based on the current speed on \code{TICK}, and
    (c) modify its speed based on \code{KEY}:

@@

The calls to \code{cb\_sim} normally happen during real-time simulation, i.e.,
while the user is interacting with the application.
However, if the peer is out of sync, the middleware may ``time travel'' and
call the simulation many times to go back and forth and catch up with real
time.
This is the reason why we define \code{START} as an event like any other:
travelling requires to simulate the execution from the beginning, event by
event.

@@

Time travelling is also the reason why \code{cb\_sim} must be separated from
the rendering callback \code{cb\_ren}~\cite{tml.js}, which would otherwise
render the screen multiple times while travelling.
%, which the middleware calls only on real-time frames.
In our example, \code{cb\_ren} just needs to redraw the rectangle at the
current position:

{\footnotesize
\begin{verbatim}
void cb_ren (void) {
    <...> // clears the SDL window
    // redraws the rectangle at the current position
    SDL_Rect r = { G.x, G.y, 20, 20 };
    SDL_DrawRect(&r);
}
\end{verbatim}
}

@@

The callback \code{cb\_evt} is called by the middleware in real time, whenever
a local SDL event occurs.
This callback has two goals:
    (a) map and broadcast application events, and
    (b) provide instant feedback to the user.
%As mentioned previously,
Not all local low-level events need to broadcast application events in the
network.
The callback is allowed to perform side effects, such as modifying the network
topology, or exhibiting instant feedback on the screen.
In our example, we only generate application events for the 5 key events, and
we also exhibit the pressed key in real time as a visual feedback:

@@

Regarding (a), if a peer experiences a time travel, it means that it received
an event after its deadline.
We count all time travel occurrences during the simulation.
%
Regarding (b), we use an absolute simulation clock to track, in real time, the
execution pace of each peer in comparison to the reference peer, which is the
one most advanced in time.
%
Regarding (c), we make peers leave and join the network periodically, possibly
creating partitions, and assert that they all catch up in real time.

@@

Figure~\ref{TODO} results...
- colors indicate feasibility

We now detail how we instrument the middleware implementation to measure the
properties above:
%
\begin{enumerate}
%
\item \textbf{Recurrence of time travels:}
Every time a peer is ahead of time and needs to travel back, we output its id
and how much ticks it needs to travel back.
At the end, we measure the percentage of the sum of all time travel ticks over
the total simulation ticks (\code{TODO} in Figure~\ref{TODO}).
As an example, if all peers sum $1000$ time travel ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
\item \textbf{Real-time pace of peers:}
Every time a peer receives \code{SYNC} packets and is behind, we output its id
and how much ticks it needs to fast forward.
At the end, we measure the percentage of the sum of all forward ticks over the
total simulation ticks (\code{TODO} in Figure~\ref{TODO}).
As an example, if all peers sum $1000$ forward ticks over $100k$ accumulated
simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a
synchronization mismatch from the beginning.
%
Each peer outputs its current tick continuously, allowing us to compare them
in the shared timeline.
Each time we see a greater tick never seen in the timeline before, we count
smaller ticks from other peers (considering latency), which would constitute
time violations (\code{TODO} in Figure~\ref{TODO}).
We expect this counter to be close to zero even considering the whole
simulation.
%
\item \textbf{Correctness of unstable peers:}
We run a separate simulation in which, for every $100s$, each peer remains
$40s$ online and $20s$ offline, including peers \code{9--11} in the middle of
the network.
This creates partitions in which sides advance in time and generate events out
of sync, which need to be merged and resynchronized constantly.
We measure
    - state (even for stable peers)
    - qto tempo leva para sync
%
\end{enumerate}

- ate 20
- 30: bugs
- 40: -> <- O(n), [500|500] future, (-5s) delay links

- qual a natureza do baksS alto e baksI baixo?
